---
title: 'R Mini Project - Covid World Vaccination Progress'
output:
  html_document:
    df_print: paged
  html_notebook:
    code_folding: none
  pdf_document: default
editor_options:
  markdown:
    wrap: 10
---

----------

**Importing
Libraries
and The
data set**

We are
using
realtime
Kaggle /
Github
which gets
updated
daily

Github:
<https://github.com/owid/covid-19-data/tree/master/public/data/vaccinations>

Kaggle:
<https://www.kaggle.com/gpreda/covid-world-vaccination-progress?select=country_vaccinations.csv>

```{r message=FALSE, warning=FALSE}
packages <- c('dplyr','ggplot2','tidyverse','countrycode','IRdisplay','htmlwidgets','plotly','tidyr','choroplethr','choroplethrMaps','lubridate','choroplethr','readr','skimr','naniar','ggmap','RColorBrewer','rworldmap','reshape2')

install_package <- packages %in% row.names(installed.packages())
if (any (install_package ==  FALSE))
  install.packages(packages[!install_package])
library("corrplot")
library("caret")
library("caTools")
sapply(packages, require, character.only = TRUE)
```

**DATA
GATHERING**

```{r message=FALSE, warning=FALSE}
data <- read_csv("./input/vaccinations.csv")
head(data)
```

**Inspect
the
structure
of data**

```{r}
#Inspect the structure of data
str(data)
```

**Inspect
the
dimensions
of the
data**

```{r}
#Inspect the dimensions of the data
dim (data)
```

```{r}
colnames(data)
```

**Counting
Total NA
in the
entire
data set**

```{r}
# Total NA in the entire data set
data1 <- data
data_vaccine <- data
paste("Total NA count:: ",sum(is.na(data_vaccine)))
```

**% of
NA's in
entire
dataset**

```{r}
# % of NA's in entire dataset
((sum(is.na(data_vaccine)) / prod(dim(data_vaccine))) * 100) %>%
round (2)
```

**NA's per
feature
can also
be
visualised
as below**

```{r}
vis_miss(data_vaccine)
```

**Total
Number of
NAs per
Feature**

```{r}
  data.frame("Total_NA" = colSums(is.na(data_vaccine))) %>%
    mutate ("Percentage_of_NA" = (colSums(is.na(data_vaccine))/dim(data_vaccine)[1]) %>% 
            round (3) * 100)
```

**Countries
for which
the data
is
collected**

```{r}
# Countries for which the data is collected

location <- unique(data_vaccine$location)
location
```

**Analysing
UK and its
Sovient
Countries**

As we
analysed
our
dataset we
got to
know that
there is
UK and
their
sovient
countries

```{r}
# There is United kingdom and also its sovient countries included in the data, 
# let us verify if they have same number of records included

 # There is United kingdom and also its sovient countries included in the data, 
# let us verify if they have same number of records included
data_vaccine %>%
filter (location %in% c('United Kingdom','Northern Ireland','Wales','Scotland'))%>%
group_by(location)%>%
count()


```

**Data
Cleaning**

```{r}
#Data Cleaning
#There are few steps in this section:

#As shown by .head() method, our data obviously contains missing values, so we'll need to deal with it somehow
#Some of the country names appeared in the dataset, does not represent a single country (constituent countries) and also the union territories
#Multiple vaccines are updated in a single row and needs to be separated

# Handling the missing values

data_vaccine[is.na(data_vaccine)] = 0

vis_miss(data_vaccine)

# there are no more NA in the data set
```

**Removing
10
countries
( UK
soviet +
UK
pertinence
+ Denmark
pertinence)**

```{r}
# Removing the UK soviet countries
# I also consider to eliminate Gibraltar, Isle of Man, Cayman Islands, Falkland Islands, Guernsey, 
# Saint Helena, Turks and Caicos Islands (UK pertinence), Faeroe Islands (Denmark pertinence).
remove_countries = c('England','Northern Ireland','Scotland','Wales','Falkland Islands','Faeroe Islands','Isle of Man','Cayman Islands','Saint Helena','Saint Lucia','Saint Vincent and the Grenadines','Saint Kitts and Nevis')

data_vaccine <-data_vaccine %>%
  filter (!location %in% remove_countries) 

unique(data_vaccine$location)
```

**Vaccines
Used
Worldwide**

```{r}
# Below the Vaccines use world wide
    data_vaccine$vaccines <- str_replace_all(data_vaccine$vaccines, " ","")
    # remove all spaces in between
    vaccine_val<- unique(data_vaccine$vaccines)
    vaccine<- vector()

    for (i in vaccine_val){
        for (j in strsplit(i, ",")){
            vaccine<- c(vaccine, j)
        }
    }
    vaccine_used<- unique(vaccine)
    vaccine_used
```

```{r}
# Inspect the usage of vaccine used wordwide

vaccine_data_val <- data.frame(matrix(ncol = length(vaccine_used), nrow = 0))
for (i in data_vaccine$vaccines){
  vaccine_data_val<- rbind(vaccine_data_val, Vectorize(grepl, USE.NAMES = TRUE)(vaccine_used, str_replace_all(i," ","")))
}
vaccine_data_val[vaccine_data_val == TRUE] = 1
vaccine_data_val[vaccine_data_val == FALSE] =0
colnames(vaccine_data_val) <- paste0(unique(vaccine))
```

**Percentage
of
Vaccines
used
Worldwide**

```{r}
vaccine_data_val %>%
summarise_all(sum)%>%
gather(key="Vaccine_name", value = "Vaccine_count")%>%
mutate (Vaccine_count1 = round(Vaccine_count/sum(Vaccine_count),3))%>%
ggplot(mapping=aes(x=reorder(Vaccine_name,-Vaccine_count1), y=Vaccine_count1, fill =Vaccine_count1, alpha=0.7))+
geom_col()+
labs(x = "Vaccines", y = "Proportions", title  = "Percentage of vaccines used worldwide")+
geom_text(aes(label = paste(Vaccine_count1*100,"%")), vjust=-0.5)+
scale_y_continuous(labels = scales::percent) + 
theme_bw()+
theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "None")
```

**Number
of
Countries
Using
these
Vaccines**

```{r}
# No of countries using these vaccines



vaccine_in_countries<- vaccine_data_val %>%
mutate(country = data_vaccine$location)%>%
group_by(country)%>%
summarise_all(sum)

data <- data.frame("No_of_countries"= apply(vaccine_in_countries[-1],2, function(c)sum(c!=0)))
cbind("Vaccine"=row.names(data),data) %>%
ggplot(mapping=aes(x=reorder(Vaccine, -No_of_countries), y=No_of_countries, fill = Vaccine, alpha=0.5))+
geom_col() +
labs(x = "Vaccines", y = "No. of Countries", title  = "Number of Countries using vaccine")+
geom_text(aes(label = No_of_countries), vjust=-0.5)+
theme_bw()+
theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "None")
```

**Spread
of Vaccine
across the
world**

```{r}
#Spread of Vaccine accrosss the world


#Spread of Vaccine accrosss the world

head(data_vaccine)


```

```{r}

world_data<- data_vaccine[,c(1,2,length(data_vaccine))] %>%
distinct()%>%
separate_rows(vaccines, sep=",")
world_data
```

**Plotting
Countries
using a
particular
vaccine**

```{r}
options(repr.plot.width=20, repr.plot.height=8)

plot_world_map <- function(iso,vacc) {
  d <- joinCountryData2Map(iso, joinCode = "ISO3",
      nameJoinColumn = "iso_code", verbose= FALSE)
  mapCountryData(d, nameColumnToPlot="vaccines", catMethod = "categorical",
    missingCountryCol = "white", addLegend = FALSE, mapTitle = paste("Countries using", vacc), oceanCol="lightblue")
    identifyCountries(d, nameCountryColumn = "iso_code",plotSelected = TRUE)
}

for (vac in vaccine_used){
    plot_world_map(world_data[world_data$vaccines == vac,],vac)
}
```

**Total
vaccination
in the
countries**

```{r}

data_vaccine <- data_vaccine %>%
  select(location, iso_code, date, total_vaccinations,people_vaccinated_per_hundred, daily_vaccinations) %>%
  group_by(location) %>%
  filter(!is.na(total_vaccinations)) %>%
  filter(total_vaccinations == max(total_vaccinations))


```

**Recent
Update
made in
the
Dataset**

```{r}
d <-as.Date(data_vaccine$date)
recent.date <- max(d)
recent.date
```

# **Top 20 Most Vaccinated Countries**

Plot of
total
vaccinations
for the
top 20
most
vaccinated
countries
as of the
most
recent
update.

```{r}
df_total_vaccinations <-data_vaccine[!(data_vaccine$location=="World"),]
df_total_vaccinations 
```

**Total
Vaccinations
per
Country**

```{r}

df_total_vaccinations %>%
  arrange(desc(total_vaccinations)) %>%
head(20)%>%
  ggplot(aes(total_vaccinations, reorder(location, total_vaccinations))) + theme_minimal() + 
  geom_bar(stat = 'identity', aes(fill=total_vaccinations)) + scale_x_continuous(labels = scales::comma) + 
  labs(title = paste('Total Vaccinations per Country as of', recent.date, sep = " "), y = '', x = 'Total Vaccinations') +
  theme(legend.position = 'none')
```

**Let's
take a
look at
vaccinations
per
hundred
people to
see how
the
percentages
look.**

```{r}
data_vaccine$people_vaccinated_per_hundred
```

**Total
Vaccination
Percentage
Per
Country**

```{r}


df_vacc_perc <- data_vaccine %>%
  select(location, iso_code, date, people_vaccinated_per_hundred) %>%
  group_by(location) %>%
  filter(!is.na(people_vaccinated_per_hundred)) %>%
  filter(people_vaccinated_per_hundred == max(people_vaccinated_per_hundred)) %>%
  distinct(location, .keep_all = TRUE)

#Plot of total vaccinations per hundred people for top 20 most vaccinated countries as of most recent update.
df_vacc_perc %>%
  arrange(desc(people_vaccinated_per_hundred)) %>%
  head(20) %>%
  ggplot(aes(people_vaccinated_per_hundred, reorder(location, people_vaccinated_per_hundred))) + theme_minimal() + 
  geom_col(aes(fill=people_vaccinated_per_hundred)) + geom_label(aes(label = people_vaccinated_per_hundred)) +
  labs(title = paste('Total Vaccination Percentage per Country as of', recent.date, sep = " "), y = '', x = 'Percentage of Population') +
  theme(legend.position = 'none')
```

# **Average Daily Vaccinations**

Let's see
how the
top
countries
are doing
as far as
daily
vaccinations.

```{r}
#Daily Vaccinations
df_vacc_daily <- data_vaccine %>%
  select(location, daily_vaccinations) %>%
  group_by(location) %>%
  filter(!is.na(daily_vaccinations))

#Aggregate data by country name and average them.
df_vacc_daily <- aggregate(list(daily_vaccinations = df_vacc_daily$daily_vaccinations), list(location = df_vacc_daily$location), mean)

#Plot of average daily vaccinations for each country.
df_vacc_daily %>%
  arrange(desc(daily_vaccinations)) %>%
  head(10) %>%
  ggplot(aes(daily_vaccinations, reorder(location, daily_vaccinations))) + theme_minimal() + 
  geom_col(aes(fill=daily_vaccinations)) + geom_label(aes(label = round(daily_vaccinations))) +
  labs(title = paste('Average Daily Vaccinations per Country as of', recent.date, sep = " "), y = '', x = 'Daily Vaccinations') +
  theme(legend.position = 'none')

```

**Feature
Scaling
for
India**

```{r}

data_India <- data1%>% filter(data1$location  == "India")
data_India

```

```{r}
data_India[is.na(data_India)] = 0
data_India
```

**Summary**

```{r}
summary(data_India)
```

```{r}
normalize <- function(x) {
  return((x-min(x))/ (max(x)-min(x)))
}
```

```{r}
normalize(data_India$people_vaccinated  )
```

```{r}
normalize(data_India$people_fully_vaccinated )
```

```{r}
normalize(data_India$daily_vaccinations_raw )
```

```{r}
normalize(data_India$ daily_vaccinations )
```

```{r}
data_India$total_vaccinations<-normalize(data_India$ total_vaccinations )
data_India$total_vaccinations
```

```{r}
cor_vaccine <- cor(data_India[, 4:11])
cor_vaccine

corrplot(cor_vaccine, method ='square', type='upper',number.cex=10)
```

**Linear
Regression**

```{r}

set.seed(123)
split<- sample.split(data_India$total_vaccinations, SplitRatio = 0.7)
train.data<- subset(data_India, split == T)
test.data <- subset(data_India, split == F)
```

```{r}
ggplot(train.data, aes(date, total_vaccinations) ) +
  geom_point() +
  stat_smooth()
```

```{r}
# Build the model
model <- lm(total_vaccinations ~ date, data = train.data)
# Make predictions
predictions <- model %>% predict(test.data)
# Model performance
data.frame(
  RMSE = RMSE(predictions, test.data$total_vaccinations),
  R2 = R2(predictions, test.data$total_vaccinations)
)
summary(model)
```

**Plotting
Linear
Regression**

```{r}
ggplot(train.data, aes(date, total_vaccinations) ) +
  geom_point() +
  stat_smooth(method = lm, formula = y ~ x)
```

**Polynomial
Regression**

```{r}
 #Build the model
class(data_India$date)
train.data$date<-as.numeric(as.Date(train.data$date, origin = "1900-01-01"))
test.data$date<-as.numeric(as.Date(test.data$date, origin = "1900-01-01"))
model2 <- lm(total_vaccinations ~ poly(date, 5, raw = TRUE), data = train.data)
# Make predictions
predictions <- model2 %>% predict(test.data)
# Model performance
data.frame(
  RMSE = RMSE(predictions, test.data$total_vaccinations),
  R2 = R2(predictions, test.data$total_vaccinations)
)
summary(model2)
```

**Plotting
Polynomial
Regression**

```{r}
ggplot(train.data, aes(date, total_vaccinations) ) +
  geom_point() +
  stat_smooth(method = lm, formula = y ~ poly(x, 5, raw = TRUE))

```
